---
title: 'Part 0: Model Generation'
author: "Max Chis"
date: "11/25/2021"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(eval = FALSE, include = FALSE)
library(tidyverse)
library(caret)
library(here)
library(Formula)
```

## Introduction

This section generates and saves the models used in other sections of the Project. 

This was done to save time when loading up the other .Rmd files, as some of these models take extended periods of time (greater than 10 minutes) to generate, and re-generating them every time I reopen RStudio strikes me as an inauspicious use of my time. 

These are saved via saveRDS to the "models" folder.

This shows the code used to generate each model, and not the output resulting from these generations, as in some cases these outputs became very verbose.

Please consult the table of contents to the left for easy navigation.

# Model Generation Setup

## Data preprocessing

Read datasets
```{r, eval=TRUE, include=TRUE}
df_start <- readr::read_csv('small_train_data.csv', col_names = TRUE)
train_x <- readr::read_csv("train_input_set_x.csv", col_names = TRUE)
train_v <- readr::read_csv("train_input_set_v.csv", col_names = TRUE)
train_outputs <- readr::read_csv("train_outputs.csv", col_names = TRUE)
```

Join train_x with train_outputs's response, and do the same with train_v 
```{r, eval=TRUE, include=TRUE}
train_x_r <- train_x %>% left_join(train_outputs, by='run_id') %>% select(!c(run_id, outcome))
train_v_r <- train_v %>% left_join(train_outputs, by='run_id') %>% select(!c(run_id, outcome))
```

Join train_x with train_output's outcome, made binary, and do the same with train_v
```{r, eval=TRUE, include=TRUE}
train_outputs$b_output <- as.factor(train_outputs$outcome)
train_x_o <- train_x %>% left_join(train_outputs, by='run_id') %>% select(!c(run_id, response, outcome))
train_v_o <- train_v %>% left_join(train_outputs, by='run_id') %>% select(!c(run_id, response, outcome))
```

## Custom Functions

Function for quickly seeding and training multiple models at once.
```{r, eval=TRUE, include=TRUE}
trainModel <- function(method_str, form, metric_str, dta, ctrl, ...) {
  set.seed(359)
  print(paste("Training model", method_str, sep=" "))
  train(form, 
        data = dta,
        method = method_str,
        metric = metric_str,
        preProcess = c("center", "scale"),
        trControl = ctrl,
        # ...
        trace = F,
        linout = F,
        importance = 'impurity'
        ) #importance needed only for ranger
}
```

```{r, eval=TRUE, include=TRUE}
run_name_save_multi_model <- function(filename, model_names, formula, metric, data, ctrl) {
  m <- model_names %>% map(trainModel, form=formula, metric_str=metric, dta=data, ctrl=ctrl)
  names(m) <- model_names
  m %>% saveRDS(here("models", paste(filename, "Rds", sep=".")))
}
```



## Constants

A set of 6 models used repeatedly in part iii and part iv
```{r, eval=TRUE, include=TRUE}
mod_names <- c('glmnet',   #Elastic Net
  'nnet',     #Neural Network
  'ranger',   #Random Forest
  'xgbTree',  #Gradient Boosted Tree
  'svmPoly',  #Support Vector Machine with Polynomial Kernel
  'kknn') #K-Nearest Neighbor
```

## Formulae

Formula for spline of x09 and x11, response
```{r, eval=TRUE, include=TRUE}
form_spline_x09_x11 <- Formula(response ~ splines::ns(x09, df=5)*splines::ns(x11, df=5))
```

Formula for spline of x09 and x11, binary output
```{r, eval=TRUE, include=TRUE}
form_spline_x09_x11_b <- Formula(b_output ~ splines::ns(x09, df=5)*splines::ns(x11, df=5))
```

Formula for interactions, triplets, of spline of v01, v06, v10, response
```{r, eval=TRUE, include=TRUE}
form_spline_v01v06v10_r <- Formula(response ~ splines::ns(v01, df=5)*splines::ns(v06, df=5)*splines::ns(v10, df=5))
```

Formula for interactions, triplets, of spline of v01,v06,v10, binary output
```{r, eval=TRUE, include=TRUE}
form_spline_v01v06v10_b <- Formula(b_output ~ splines::ns(v01, df=5)*splines::ns(v06, df=5)*splines::ns(v10, df=5))
```

Formula for spline of v10 in interaction with the additive inputs of v01, v06, v08, response
```{r, eval=TRUE, include=TRUE}
form_spline_v10_int_v01v06_v08_r <- Formula(response ~ splines::ns(v10, df=5)*(v01+v06+v08))
```

Formula for spline of v10 in interaction with the additive inputs of v01, v06, v08, binary output
```{r, eval=TRUE, include=TRUE}
form_spline_v10_int_v01v06_v08_b <- Formula(b_output ~ splines::ns(v10, df=5)*(v01+v06+v08))
```

Formula for spline of pairs of v01, v06, v10, response
```{r}
form_v_6_models_spline_pairs_v01v06v10_r <- Formula(response ~ (splines::ns(v01, df=5) + splines::ns(v06, df=5) + splines::ns(v10, df=5))^2)
```

Formula for spline of pairs of v01, v06, v10, binary output
```{r}
form_v_6_models_spline_pairs_v01v06v10_b <- Formula(b_output ~ (splines::ns(v01, df=5) + splines::ns(v06, df=5) + splines::ns(v10, df=5))^2)
```

## Validation 

Validation using simple 5-fold cross-validation, mainly due to speed concerns.
```{r, eval=TRUE, include=TRUE}
ctrl <- trainControl(
  method='cv',
  number = 5,
  verboseIter = TRUE
)
```

Control for calculating ROC
```{r, eval=TRUE, include=TRUE}
ctrl_ROC <- trainControl(
  method='cv',
  number=5,
  verboseIter=T,
  classProbs=T,
  summaryFunction = twoClassSummary
)
```


# Part 2

## Linear Models

### mod_start_additive_all
All 5 starter inputs with additive features only
```{r eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ ., data=df_start) %>%
saveRDS(here("models", "mod_start_additive_all.Rds"))
```

### mod_start_pairwise_all
All pair-wise interactions with start inputs

```{r eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ .^2, data=df_start) %>%
  saveRDS(here("models", "mod_start_pairwise_all.Rds"))
```

### mod_start_2nd_poly_all
Polynomial basis of 2nd order applied to all 5 inputs

```{r eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ (
    x07 + x09 + x10 + x11 + x21 + I(x07^2) + I(x09^2) + I(x10^2) + I(x11^2) + I(x21^2)
  ), data=df_start) %>% 
  saveRDS(here("models", "mod_start_2nd_poly_all.Rds"))
```

### mod_start_lq_ap_x9_x11
Linear and quadratic additive with all-pairwise interactions for starter x09 and x11.

```{r eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ (x09+I(x09^2))*(x11+I(x11^2)), data = df_start) %>%
  saveRDS(here("models", "mod_start_lq_ap_x09_x11.Rds"))
```

### mod_start_spline_x09_x11
Spline basis for starter x09 and x11

```{r eval=FALSE, include=TRUE}
set.seed(359)
mod_spline_x09_x11 <- lm(form_spline_x09_x11, data = df_start) %>%
  saveRDS(here("models", "mod_start_spline_x09_x11.Rds"))
```

### mod_start_interact_x9_x11
Interaction of starter x09 and x11 as well as x09 and x11 themselves.
```{r eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ (x09*x11), data = df_start) %>%
  saveRDS(here("models", "mod_start_interact_x09_x11.Rds"))
```

## Bayesian model (Stan_lm)

### mod_start_bayes_lq_ap_x9_x11
Bayesian linear model for linear/quadratic additive with all-pairwise interactions for starter x09 and x11
```{r eval=FALSE, include=TRUE}
set.seed(359)
stan_lm(response ~ (x09+I(x09^2))*(x11+I(x11^2)), 
                 data = df_start,
                 prior = NULL,
                 seed = 432123) %>% 
  saveRDS(here("models", "mod_start_bayes_lq_ap_x09_x11.Rds"))
```
Note: Received warning of "4 divergent transitions after warmup" upon running this.

### mod_start_bayes_spline_x09_x11
Bayesian linear model spline basis for starter x09 and x11
```{r eval=FALSE, include=TRUE}
set.seed(359)
stan_lm(form_spline_x09_x11, data = df_start,
                 prior = NULL,
                 seed = 432123) %>% 
  saveRDS(here("models", "mod_start_bayes_spline_x09_x11.Rds"))
```


# Part 3

## X-Variable Models

### mod_x_additive_all
All x inputs with additive features
```{r, eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ ., data=train_x_r) %>% 
  saveRDS(here("models", "mod_x_additive_all.Rds"))
```

### mod_x_pairwise_all
All x inputs and interactions
```{r, eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ .^2, data=train_x_r) %>% 
  saveRDS(here("models", "mod_x_pairwise_all.Rds"))
```

### mod_x_spline_x09_x11
Spline models for x09 and x11
```{r, eval=FALSE, include=TRUE}
set.seed(359)
lm(form_spline_x09_x11, data = train_x_r) %>% 
  saveRDS(here("models", "mod_x_spline_x09_x11.Rds"))
```

### mod_x_enet_pairwise_all
Regularized regression with Elastic net: All pair-wise interactions between x inputs
```{r, eval=FALSE, include=TRUE}
trainModel(method="glmnet", form=response ~ .^2, metric_str='RMSE', dta=train_x_r, ctrl=ctrl) %>%
  saveRDS(here("models", "mod_x_enet_pairwise_all.Rds"))
```
Using proc.time() to time myself, this took roughly 1 minute to train.

### mod_x_6_models_spline_x09_x11_RMSE

Six different models for the spline of x09 and x11:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor
```{r message=FALSE, warning=FALSE, results = 'hide', include=TRUE, eval=FALSE}
run_name_save_multi_model(filename="mod_x_6_models_spline_x09_x11_RMSE", model_names=mod_names, formula=form_spline_x09_x11, metric='RMSE', data=train_x_r, ctrl=ctrl)
```




## V-Variable Models

### mod_v_additive_all
All v inputs with additive features
```{r, eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ ., data=train_v_r) %>% saveRDS(here("models", "mod_v_additive_all.Rds"))
```

### mod_v_pairwise_all

All v inputs with all pairwise interactions
```{r, eval=FALSE, include=TRUE}
set.seed(359)
lm(response ~ .^2, data=train_v_r) %>% saveRDS(here("models", "mod_v_pairwise_all.Rds"))
```

### mod_v_enet_pairwise_all
Regularized regression with Elastic net: All pair-wise interactions between v inputs
```{r eval=FALSE, include=TRUE}
trainModel(method="glmnet", form=response ~ .^2, metric_str='RMSE', dta=train_v_r, ctrl=ctrl) %>% 
  saveRDS(here("models", "mod_v_enet_pairwise_all.Rds"))
```
Using proc.time() to time myself, this took roughly 5 minutes to train.

### mod_v_6_models_spline_v01v06v10_RMSE
Six different models for all pairwise and triplet interactions of the splines of v01, v06, and v10:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor
```{r eval=FALSE, include=TRUE}
run_name_save_multi_model(filename="mod_v_6_models_spline_v01v06v10_RMSE", model_names=mod_names, formula=form_spline_v01v06v10_r, metric='RMSE', data=train_v_r, ctrl=ctrl)
```
Note: Do not, for the love of God, run this again. It took several hours.

### mod_v_6_models_spline_pairs_v01v06v10_r
Six different models for all pairwise interactions of the splines of v01, v06, and v10:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor
```{r}
run_name_save_multi_model(filename="mod_v_6_models_spline_pairs_v01v06v10_r", model_names=mod_names, formula=form_v_6_models_spline_pairs_v01v06v10_r, metric='RMSE', data=train_v_r, ctrl=ctrl)
```

# Part 4

## X-Variables

### mod_x_additive_all_b
All x inputs with additive features, with response value as binary output
```{r eval=FALSE, include=FALSE}
set.seed(359)
glm(b_output ~ ., family = "binomial", data = train_x_o) %>% 
  saveRDS(here("models","mod_x_additive_all_b.Rds"))
```

### mod_x_pairwise_all_b
All x inputs with all pairwise interactions, with response value as binary output.
```{r eval=FALSE, include=TRUE}
ptm <- proc.time()
m <- glm(b_output ~ .^2, family = "binomial", data = train_x_o) 
m %>% saveRDS(here("models","mod_v_pairwise_all_b.Rds"))
proc.time() - ptm
```
Results did not converge.
Using proc.time() to time myself, this took 48 seconds to train.

### mod_x_glmnet_pairwise_all_b

## Accuracy
Glmnet model of all x inputs with all pairwise interactions, with response value as binary input, modeled to maximize Accuracy.
```{r eval=FALSE, include=TRUE}
m <- trainModel(method="glmnet", form=b_output ~ .^2, metric_str='Accuracy', dta=train_x_o, ctrl=ctrl)
m %>% saveRDS(here("models", "mod_x_glmnet_pairwise_all_b_Acc.Rds"))
```

## ROC/AUC
Glmnet model of all x inputs with all pairwise interactions, with response value as binary input, modeled to maximize ROC/AUC.
```{r eval=FALSE, include=TRUE}
m <- trainModel(method="glmnet", form=b_output ~ .^2, metric_str='ROC', dta=train_x_o, ctrl=ctrl_ROC)
m %>% saveRDS(here("models", "mod_x_glmnet_pairwise_all_b_ROC.Rds"))
```

### mod_x_6_models_spline_x09_x11_b

####AUC/ROC
Six different models for all pairwise interactions of the splines of x09 and x11,with response value as binary input, modeled to maximize ROC:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor


```{r eval=FALSE, include=TRUE}
run_name_save_multi_model(filename="mod_x_6_models_spline_x09_x11_b_AUC", model_names=mod_names, formula=form_spline_x09_x11_b, metric='ROC', data=train_x_o, ctrl=ctrl_ROC)
```

#### Accuracy
Six different models for all pairwise interactions of the splines of x09 and x11,with response value as binary input, modeled to maximize Accuracy:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor

```{r message=FALSE, warning=FALSE, results = 'hide', include=TRUE, eval=FALSE}
run_name_save_multi_model(filename="mod_x_6_models_spline_x09_x11_b_Acc", model_names=mod_names, formula=form_spline_x09_x11_b, metric='Accuracy', data=train_x_o, ctrl=ctrl)
```

Time required to model: 280 seconds.

## V-Variables

### mod_v_additive_all_b
All v inputs with additive features, with response value as binary output
```{r eval=FALSE, include=TRUE}
set.seed(359)
m <- glm(b_output ~ ., family = "binomial", data = train_v_o) 
m %>% saveRDS(here("models","mod_v_additive_all_b.Rds"))
```


### mod_v_pairwise_all_b
All v inputs with additive features, with response value as binary output
```{r eval=FALSE, include=TRUE}
ptm <- proc.time()
set.seed(359)
m <- glm(b_output ~ .^2, family = "binomial", data = train_v_o) 
m %>% saveRDS(here("models","mod_v_pairwise_all_b.Rds"))
proc.time() - ptm
```
Result did not converge.
Time required to model: 47 seconds.

### mod_v_glmnet_pairwise_all_b

#### Accuracy
Glmnet model of all v inputs with all pairwise interactions, with response value as binary input, modeled to maximize Accuracy.
```{r eval=FALSE, include=TRUE}
m <- trainModel(method="glmnet", form=b_output ~ .^2, metric_str='Accuracy', dta=train_v_o, ctrl=ctrl)
m %>% saveRDS(here("models", "mod_v_glmnet_pairwise_all_b_Acc.Rds"))
```

#### ROC/AUC
Glmnet model of all v inputs with all pairwise interactions, with response value as binary input, modeled to maximize ROC/AUC
```{r eval=FALSE, include=TRUE}
m <- trainModel(method="glmnet", form=b_output ~ .^2, metric_str='ROC', dta=train_v_o, ctrl=ctrl_ROC)
m %>% saveRDS(here("models", "mod_v_glmnet_pairwise_all_b_ROC.Rds"))
```


### mod_v_6_models_spline_pairs_v01v06v10

#### Accuracy
Six different models for all pairwise interactions of the splines of v01, v06 and v10, with response value as binary input, modeled to maximize Accuracy:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor
```{r}
run_name_save_multi_model(filename="mod_v_6_models_spline_pairs_v01v06v10_Accuracy", model_names=mod_names, formula=form_v_6_models_spline_pairs_v01v06v10_b, metric='Accuracy', data=train_v_o, ctrl=ctrl)
```

#### ROC
Six different models for all pairwise interactions of the splines of v01, v06 and v10, with response value as binary input, modeled to maximize ROC:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor

```{r}
run_name_save_multi_model(filename="mod_v_6_models_spline_pairs_v01v06v10_ROC", model_names=mod_names, formula=form_v_6_models_spline_pairs_v01v06v10_b, metric='ROC', data=train_v_o, ctrl=ctrl_ROC)
```
