---
title: "Rejected Models"
author: "Max Chis"
date: "12/1/2021"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### mod_v_6_models_spline_v01_v02_RMSE

Interaction of all splines of v01 and v02.
Rejected due to high RMSE (in range of 0.60-0.70)

Formula for interactions, pairwise, of spline of v01 and v02, response
```{r, eval=TRUE, include=TRUE}
form_spline_v01_v02 <- Formula(response ~ splines::ns(v01, df=5)*splines::ns(v02, df=5))
```

Formula for interactions, pairwise, of spline of v01 and v02, binary output
```{r, eval=TRUE, include=TRUE}
form_spline_v01_v02_b <- Formula(b_output ~ splines::ns(v01, df=5)*splines::ns(v02, df=5))
```

Six different models for the spline of v01 and v02:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor

Generate RMSE Model
```{r eval=FALSE, include=TRUE}
run_name_save_multi_model(filename="mod_v_6_models_spline_v01_v02_RMSE", model_names=mod_names, formula=form_spline_v01_v02, metric='RMSE', data=train_v_r, ctrl=ctrl)
```
Using proc.time() to time myself, this took roughly 5 minutes to train.

Generate Accuracy Model
```{r}
run_name_save_multi_model(filename="mod_v_6_models_spline_v01_v02_b_Acc", model_names=mod_names, formula=form_spline_v01_v02_b, metric='Accuracy', data=train_v_o, ctrl=ctrl)
```

Generate ROC Model
```{r}
run_name_save_multi_model(filename="mod_v_6_models_spline_v01_v02_b_ROC", model_names=mod_names, formula=form_spline_v01_v02_b, metric='ROC', data=train_v_o, ctrl=ctrl_ROC)
```

Read in file
```{r}
mod_v_6_models_spline_v01_v02_RMSE <- readRDS(here("models", "mod_v_6_models_spline_v01_v02_RMSE.Rds"))
```

Dotplot 
```{r}
all_fit_metrics <- mod_v_6_models_spline_v01_v02_RMSE %>% resamples()
dotplot(all_fit_metrics, metric = "RMSE")
```

### mod_v_6_models_spline_v10_int_v01v06_v08_r
Six different models for the pairwise interactions of all splines of v10 with the variables of v01, v06, and v08
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor
```{r message=FALSE, warning=FALSE, results = 'hide', include=TRUE, eval=FALSE}
run_name_save_multi_model(filename="mod_v_6_models_spline_v10_int_v01v06_v08_r", model_names=mod_names, formula=form_spline_v10_int_v01v06_v08_r, metric='RMSE', data=train_v_r, ctrl=ctrl)
```

#### RMSE 
Rejected due to insufficiently distinct performance compared with other models
```{r}
mod_v_6_models_spline_v10_int_v01v06_v08_r <- readRDS(here("models", "mod_v_6_models_spline_v10_int_v01v06_v08_r.Rds"))
mod_v_6_models_spline_v10_int_v01v06_v08_r %>% resamples() %>% dotplot(metric="RMSE", main="spline_v10_int_v01v06_v08_r")
mod_v_6_models_spline_v10_int_v01v06_v08_r$glmnet$finalModel %>% coefplot::coefplot()
```

### mod_v_6_models_spline_v01v06v10
Six different models for all pairwise and triplet interactions of the splines of v01, v06 and v10. Rejected due to performance insufficiently distinct from simpler models, and excessive length of training time. 

#### RMSE
```{r}
mod_v_6_models_spline_v01v06v10_RMSE <- readRDS(here("models", "mod_v_6_models_spline_v01v06v10_RMSE.Rds"))
mod_v_6_models_spline_v01v06v10_RMSE %>% resamples() %>% dotplot(metric="RMSE", main="spline_v01v06v10_RMSE")
mod_v_6_models_spline_v01v06v10_RMSE$glmnet$finalModel %>% coefplot::coefplot()
```

#### Accuracy
, with response value as binary input, modeled to maximize Accuracy:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor

```{r eval=FALSE, include=TRUE}
run_name_save_multi_model(filename="mod_spline_v10_int_v01v06_v08_b_Acc", model_names=mod_names, formula=form_spline_v10_int_v01v06_v08_b, metric='Accuracy', data=train_v_o, ctrl=ctrl)
```


#### AUC
Six different models for all pairwise and triplet interactions of the splines of v01, v06 and v10, with response value as binary input, modeled to maximize ROC:
* Elastic Net
* Neural Network
* Random Forest
* Gradient Boosted Tree
* Support Vector Machine with Polynomial Kernel
* K-Nearest neighbor

```{r eval=FALSE, include=TRUE}
run_name_save_multi_model(filename="mod_spline_v10_int_v01v06_v08_b_ROC", model_names=mod_names, formula=form_spline_v10_int_v01v06_v08_b, metric='ROC', data=train_v_o, ctrl=ctrl_ROC)
```
Rejected due to inferior performance, over-complexity, and excessive length of training time.



